Forecasting 2024 U.S. Election using Polls-of-Polls

### 2024-10-23 paper/paper.qmd
fix reference and include survey copy in appendix

---
title: Forcasting 2024 U.S. Election using Polls-of-Polls
subtitle: Using a Linear Regression Model to Adjust Pollster Biases and Predict Kamala Harris' Vote Percentage at 47.78%
author: 
  - Shamayla Durrin
  - Denise Chang 
  - Krishna Kumar
thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."
date: today
date-format: long
abstract: "This paper uses a polls of polls approach to predict voter support for Kamala Harris in the 2024 U.S. election, aggregating multiple polls to reduce individual bias and improve accuracy. Through a regression model incorporating factors such as pollster, sample size, state, and recency, we predict her overall support to be 47.78%, just below the 50% threshold. Our analysis highlights the importance of poll quality and recency in understanding trends in public opinion. These findings are significant as they provide a more reliable estimate of electoral outcomes, helping to inform political strategies and public discourse."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

The raw data for this project, sourced from FiveThirtyEight, [@FiveThirtyEight] underwent a series of cleaning steps to prepare it for analysis. Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the janitor package [@janitor]. A new binary variable, 'national', was created to indicate whether a poll was conducted at the national or state level. Missing values in the 'state' column were replaced with "Not Applicable," and numeric grades were evaluated to filter out low-quality pollsters, keeping only those with a numeric grade above 1. This cutoff was selected to retain mid to high-level pollsters for more reliable results. These steps were performed using functions from the dpylr package \[\@dplyr\]. Furthermore, dates were standardized and converted into a proper format for analysis using the lubridate package \[\@lubridate\]. Polls related to Kamala Harris were retained for further analysis, and percentage support values were transformed into actual numbers of supporters based on sample size. Additionally, pollster counts below five were excluded to focus on more reliable data sources. Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the arrow package \[\@arrow\].

@@

# Forcasting Election Outcome through Pooling Polls

@@

\[\@fig-sample\] suggests that the relationship between support for Kamala Harris and sample size shows a slight positive trend, indicating a small increase in support as sample sizes grow. Both national (blue) and state (green) polls reflect this upward trend, although the effect is minimal. Since voting outcomes often depend on small margins, this subtle increase in support with larger sample sizes could still be relevant in predicting overall support.

@@

In \[\@fig-enddate\], we observe a slight increasing trend in support for Kamala Harris as the end date of polls progresses towards October. This may indicate a gradual improvement in her polling performance over time. The color gradient, reflecting the pollster's error and bias (Pollscore), reveals variations in bias: greener points represent more negative biases, while redder points indicate more positive biases.

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-mapinterac
#| fig-cap: "Average Support for Kamala Harris in Different States estimated by Polls"

@@

### Prediction

@@

A link to the survey can be found at: https://forms.gle/haGLaQBPQhrXE6Xa7 \newpage

>>>>

---
title: Forecasting 2024 U.S. Election using Polls-of-Polls
subtitle: Using a Linear Regression Model to Adjust Pollster Biases and Predict Kamala Harris Vote Percentage at 47.78%
author: 
  - Shamayla Durrin
  - Denise Chang 
  - Krishna Kumar
thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."
date: today
date-format: long
abstract: "This paper uses a polls of polls approach to predict voter support for Kamala Harris in the 2024 U.S. election, aggregating multiple polls to reduce individual bias and improve accuracy. Through a regression model incorporating factors such as pollster, sample size, state, and recency, we predict her overall support to be 47.78%, just below the 50% threshold. Our analysis highlights the importance of poll quality and recency in understanding trends in public opinion. These findings are significant as they provide a more reliable estimate of electoral outcomes, helping to inform political strategies and public discourse."
format: 
  pdf:
    documentclass: article
    number-sections: true
    header-includes:
      - \usepackage{pdfpages}
bibliography: references.bib
---

@@

The raw data for this project, sourced from FiveThirtyEight, [@FiveThirtyEight] underwent a series of cleaning steps to prepare it for analysis. Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the `janitor` package [@janitor]. A new binary variable, 'national', was created to indicate whether a poll was conducted at the national or state level. Missing values in the 'state' column were replaced with "Not Applicable," and numeric grades were evaluated to filter out low-quality pollsters, keeping only those with a numeric grade above 1. This cutoff was selected to retain mid to high-level pollsters for more reliable results. These steps were performed using functions from the `dpylr` package [@dplyr]. Furthermore, dates were standardized and converted into a proper format for analysis using the `lubridate` package [@lubridate]. Polls related to Kamala Harris were retained for further analysis, and percentage support values were transformed into actual numbers of supporters based on sample size. Additionally, pollster counts below five were excluded to focus on more reliable data sources. Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the `arrow` package [@arrow].

@@

# Forecasting Election Outcome through Pooling Polls

@@

[@fig-sample] suggests that the relationship between support for Kamala Harris and sample size shows a slight positive trend, indicating a small increase in support as sample sizes grow. Both national (blue) and state (green) polls reflect this upward trend, although the effect is minimal. Since voting outcomes often depend on small margins, this subtle increase in support with larger sample sizes could still be relevant in predicting overall support.

@@

In [@fig-enddate], we observe a slight increasing trend in support for Kamala Harris as the end date of polls progresses towards October. This may indicate a gradual improvement in her polling performance over time. The color gradient, reflecting the pollster's error and bias (Pollscore), reveals variations in bias: greener points represent more negative biases, while redder points indicate more positive biases.

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-enddate
#| fig-cap: "Average Support for Kamala Harris in Different States estimated by Polls"

@@

@@

A link to the survey can be found at: https://forms.gle/haGLaQBPQhrXE6Xa7

# Copy of the Survey
\includepdf[pages=-]{../other/surveyfiles/SurveyCopy.pdf}
